Semantic Linearity in a Lazy Optimising Compiler

### 2025-02-22 icfp.tex
2.1: finish pass + acknowledgements

\paragraph{Case expressions}
Finally, we discuss case expressions,
which are the key ingredient that brings together the
semantic linearity insights developed thus far given that
\emph{case expressions drive evaluation}, and semantic linearity can only be
understood in terms of how expressions are evaluated.

Up until now, there
have been no examples in which linear resources are \emph{fully consumed} in
the bodies of linear functions. In other words, all example functions so far
return a value that has to be itself consumed exactly once to ensure the linear
argument is, in turn, consumed exactly once -- as opposed to functions whose
application simply needs to be evaluated to guarantee its linear argument is
consumed (e.g.,~functions that return an unrestricted value).
%
%The latter are of particular relevance because linearly-typed abstractions
%usually require such a function to provide newly-created linear resources to.
%
% For example, the entry point to the linear array API presented in Linear
% Haskell takes such a function as its second argument:
% \begin{working}
% \begin{lstlisting}
% newMArray :: Int -> (MArray a *' $\lolli$ '* Unrestricted b) *' $\lolli$ '* b
% \end{lstlisting}
% \end{working}
%(The second argument is a function that consumes |MArray a| linearly
%and returns an unrestricted result -- we don't need to consume said result
%exactly once to guarantee that |MArray a| is used linearly in the function
%body.)
%
% ROMES:TODO: Há qualquer coisa que não gosto nada! na coerência deste paragrafo
%
Case expressions enable us to consume resources and so write
functions that fully consume their linear arguments.
% CORTEI
To understand exactly how,
we turn to the definition of \emph{consuming} a resource from Linear
Haskell~\cite{cite:linearhaskell}:
%begin{itemize}
% \item
to consume a value of atomic base type (such as~\texttt{Int} or
        \texttt{Ptr}) exactly once, just evaluate it;
        % \item
        to consume a function value exactly once, apply it to one argument,
        and consume its result exactly once;
        % \item
        to consume a value of an algebraic datatype exactly once,
        pattern-match on it, and consume all its linear components exactly once.
        % For example, a linear pair (equivalent to $\tensor$) is consumed exactly
        % once if pattern-matched on \emph{and} both the first and second element are
        % consumed once.
%\end{itemize}
Thus, we can consume a linear resource by fully evaluating it, through case
expressions.

@@

In Core, case expressions are of the form $\ccase{e_s}{z~\{\ov{\rho_i \to e_i}\}}$,
where $e_s$ is the case \emph{scrutinee}, $z$ is the case \emph{binder}, and
$\ov{\rho_i \to e_i}$ are the case \emph{alternatives}, composed of a pattern
$\rho_i$ and of the corresponding expression $e_i$. 
%\begin{enumerate}
%\item
The case scrutinee is always evaluated to Weak Head Normal Form (WHNF) and
%
% \item
% evaluating an expression that is already in WHNF is a no-op, that is,
% no computation occurs; 
%
the case binder is an alias to the result of evaluating the scrutinee
to WHNF. Moreover, in Core, the alternative patterns are always exhaustive.
% i.e. there always exists
% a pattern that matches the WHNF of a value resulting from evaluating the
% scrutinee, where a pattern is either a wildcard that matches all expressions
% ($\_$), or a constructor and its linear and non-linear component binders
% ($K~\ov{x}\ov{y}$, with $\ov{x}$ as linearly-bound variables and $\ov{y}$ as
% unrestricted ones).

@@

We start with an
example of a program that constructs a tuple from linear resources then pattern
matches on it, then uses both linearly-bound variables from the tuple pattern
match. This is well-typed in Linear Haskell:
\begin{working}
\begin{lstlisting}
f6 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* c) -> c
f6 x y use = case (x,y) of z { (a,b) -> use a b }
\end{lstlisting}
\end{working}
What might be more surprising is that a similar program which discards the
pattern variables and instead uses the resources in the scrutinee is 
\emph{semantically} linear, despite not being accepted by Linear Haskell:
\begin{notyet}
\begin{lstlisting}
f7 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* c) -> c
f7 x y use = case (x,y) of z { (a,b) -> use x y }
\end{lstlisting}
\end{notyet}
We justify that \incode{f7} is linear by appealing to the fact that 
the tuple being scrutinized is already in WHNF and so evaluating it will
not consume either $x$ nor $y$. Even if the tuple was constructed with
two expressions using $x$ and $y$ respectively, no computation would happen
since we are not using $a$ or $b$ (thereby never forcing the arguments
of the tuple). However, if we did use $a$ in the case body, then $x$
would be unavailable (and so \incode{f8} below cannot be accepted):
\begin{noway}
\begin{lstlisting}
f8 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* a *' $\lolli$ '* c) -> c
f8 x y use = case (x,y) of z { (a,b) -> use a x }
\end{lstlisting}
\end{noway}
This idea that $x$ and $a$ are mutually exclusive is the same behind let
bindings being mutually exclusive to the resources that define them.
By forcing the pattern variable we run the computations
defined in terms of the linear variables used for that constructor
argument. Otherwise, if we do not use the pattern variables, then we
do not run the computation and thus no resources are consumed.

A third (semantically well-typed, but rejected by Linear Haskell) option is to use the case binder $z$ instead of $a,b$ or $x,y$:
\begin{notyet}
\begin{lstlisting}
f9 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* c) -> c
f9 x y use = case (x,y) of z { (a,b) -> uncurry use z }
\end{lstlisting}
\end{notyet}
Again, $z$ is mutually exclusive with $(a,b)$ and with $(x,y)$, but at least one of
the three must occur to ensure the linear resources are consumed. In this
example, we can think that using $a$ entails using the resource $x$, $b$ the
resource $y$, and the case binder $z$ entails using both $a$ and $b$.
%
Dually, consider the scrutinee to be an expression that's not in WHNF, s.t.
evaluating it to WHNF will require performing computation and thus consume linear
resources (\incode{f10} below is well-typed in our system but rejected
by Linear Haskell):
\begin{notyet}
\begin{lstlisting}
f10 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* (c,d)) -> (c,d)
f10 x y use = case use x y of z { (a,b) -> z }
\end{lstlisting}
\end{notyet}
Unlike when the scrutinee was in WHNF, we can no longer use $x,y$ in the case
alternatives, but we must still use either the case binder $z$ or the linear
pattern variables $a,b$. %  e.g. it would be quite disastrous if any of the following typechecked:
% \begin{noway}
% \begin{lstlisting}
% doubleFree :: Ptr *' $\lolli$ '* (Ptr *' $\lolli$ '* Result) -> Result
% doubleFree x free = case free x of z { Result v -> free x }
% \end{lstlisting}
% \end{noway}
% % Remover
% \begin{noway}
% \begin{lstlisting}
% leakPointer :: Ptr -> ()
% leakPointer x = case id x of z { _ -> () }
% \end{lstlisting}
% \end{noway}
%
%whereas this is fine:
%\begin{notyet}
%\begin{code}
%f16 :: Ptr -o (Ptr -o (Value,Int)) -> (Value,Int)
%f16 x free = case K (free x) of z { K y -> free x }
%\end{code}
%\end{notyet}
%
The result of evaluating the scrutinee must be consumed exactly once to guarantee
that the resources used in the scrutinee are fully consumed. %, or
                                %risk them being only ``almost''
                                %consumed.
For instance, take
\incode{use} in \incode{f10} to be the pairing function \incode{(,)}: it is not sufficient for \incode{use x y} to be
evaluated to WHNF to consume \incode{x} and \incode{y}. Otherwise, if all the resources were
considered to be fully consumed after the scrutinee were evaluated in a case
expression, we could simply ignore the pattern variables, potentially
discarding linear resources. In
short, if the scrutinee is not in WHNF we must either consume the case binder
or the linear components of the pattern.

We now focus on pattern matching on constructors without
any linear components. If the constructor has no linear fields, it means the
result can be consumed unrestrictedly and, therefore, all linear resources used
in the computation have been fully consumed.
%
Consequently, in a branch of a constructor without linear fields we know the
result of evaluating the scrutinee to be unrestricted, so we can use the case
binder unrestrictedly and refer to it zero or more times. For example,
the program in Figure~\ref{fig:nolin} is semantically linear.
%
\begin{figure}[t]
 \begin{minipage}{0.5\textwidth}
\begin{notyet}
\begin{lstlisting}
f11 :: () *' $\lolli$ '* ()
f11 x = case x of z { () -> z <> z }
\end{lstlisting}
\end{notyet}
\vspace{-0.5cm}
\caption{No Linear Fields\label{fig:nolin}}
\end{minipage}%
\begin{minipage}{0.5\textwidth}
\begin{limitation}
\begin{lstlisting}
f12 :: a *' $\lolli$ '* a
f12 x = case K1 x of z { K2 -> x; K1 a -> x }
\end{lstlisting}
\end{limitation}
\vspace{-0.5cm}
\caption{Absurd Branches\label{fig:absurd}}
\end{minipage}
\end{figure}
%
A second example of an unrestricted pattern, where \incode{K2} has no fields and \incode{K1}
has one linear field, is shown in Figure~\ref{fig:absurd}. Seemingly, the resource
$x$ must have been fully consumed to take the \incode{K2} branch, but because the
scrutinee is known to be \incode{K1}, the \incode{K2} branch is \emph{absurd}, so, in reality
any resource could be freely used in that branch, and the example is
\emph{semantically} linear (despite not being seen as so by our system).

Many of these examples might seem unrealistic due to the pattern
match on a known constructor, but we note that, during the
transformations programs undergo in an optimising compiler, many
such programs naturally occur (e.g.,~if the definition of a
function is inlined in the scrutinee).
%
% However, the scrutinee before evaluation likely isn't in WHNF, thus scrutinee
% resources cannot directly occur in the case alternatives, since they might be
% potentially consumed when evaluating the scrutinee, the program leading up to
% this one wouldn't have been linear (even semantically).

@@

As for \emph{default} case alternatives, also known as \emph{wildcards}
(written $\_$): matching a wildcard doesn't provide new linearity information,
unlike matching against a pattern with or without linear components, however,
the scrutinee will still be evaluated to WHNF.
%
% so linearity is seen as before but without
% fully consuming the scrutinee linear resources (in non-linear patterns) nor
% binding new linear resources (in linear patterns).
%
Briefly, when matching a wildcard, if the scrutinee is already in WHNF, we can
either use the resources from the scrutinee directly, or the case binder in
that alternative. If the scrutinee is not in WHNF, we \emph{must} use the case
binder, as it's the only way to linearly consume the result of evaluating the
scrutinee.

Finally, we discuss case expressions which scrutinize a
single variable $x$, e.g.:
$
\lambda x.~\ccase{x}{\_ \to x}
$. 
It might seem as though this function is linear:
%begin{itemize}
% \item
if the argument instantiating $x$ is in WHNF, then scrutinizing it is a no-op, and returning $x$
just returns the resource intact;
% \item
if the argument instantiating $x$ is not in WHNF, then scrutinizing it evaluates it to WHNF, and returning
$x$ returns the result of evaluating $x$ that still had to be consumed exactly
once.
%\end{itemize}
However, in practice, this program's linearity can only be determined by
further analysis on the non-strict evaluation strategy. If linear
function applications, in general, are $\beta$-reduced \emph{call-by-name} (a desirable
practice, as linear functions are known to use their argument exactly once)
% and thus call-by-name reduction will definitely not duplicate computation
while the above function is considered linear, then such an application might
duplicate linear resources during evaluation. For instance:
\[
\begin{array}{l}
(\lambda x.~\ccase{x}{\_ \to x})~(free~x)
\Longrightarrow_{CBN}
\ccase{free~x}{\_ \to free~x}
\end{array}
\]
On the other hand, if linear function applications are reduced
\emph{call-by-need}, the above example would instead be semantically linear
because the result of freeing $x$ would be bound to some $y$ first.
%
The system we present in the next section, which evaluates linear
applications call-by-name, does not accept the above function scrutinizing a
variable as linear. 
This interaction between evaluation and linearity arises in the type
preservation proof of our system, and is again explored with the reverse
binder swap transformation discussed in
Section~\ref{sec:optimisations-preserve-types-meta}.

@@

>>>>

\paragraph{Case expressions}
Above, we have observed more programs can be accepted as linear by delaying
resource consumption in (recursive) let bindings until they are evaluated.
On the other hand, \emph{case expressions drive evaluation} -- and, therefore,
do consume resources.

%and semantic linearity can only be understood in terms of how expressions are
%evaluated.

%Up until now, there have been no examples in which linear resources are
%\emph{fully consumed} in the bodies of linear functions.

%%%%%%%%%%%%%%% KEEEP ????? %%%%%%%
%%%%%%%%%All example functions so far returned a value that has to be itself consumed
%%%%%%%%%exactly once to ensure the linear argument is, in turn, consumed exactly once
%%%%%%%%%-- as opposed to functions whose application simply needs to be evaluated to
%%%%%%%%%guarantee its linear argument is consumed (e.g.,~functions that return an
%%%%%%%%%unrestricted value).

%
%The latter are of particular relevance because linearly-typed abstractions
%usually require such a function to provide newly-created linear resources to.
%
% For example, the entry point to the linear array API presented in Linear
% Haskell takes such a function as its second argument:
% \begin{working}
% \begin{lstlisting}
% newMArray :: Int -> (MArray a *' $\lolli$ '* Unrestricted b) *' $\lolli$ '* b
% \end{lstlisting}
% \end{working}
%(The second argument is a function that consumes |MArray a| linearly
%and returns an unrestricted result -- we don't need to consume said result
%exactly once to guarantee that |MArray a| is used linearly in the function
%body.)
%
% ROMES:TODO: Há qualquer coisa que não gosto nada! na coerência deste paragrafo
%
%Case expressions enable us to consume resources and so write
%functions that fully consume their linear arguments.
% CORTEI
To understand exactly how,
we turn to the definition of \emph{consuming} a resource from Linear
Haskell~\cite{cite:linearhaskell}:
%begin{itemize}
% \item
to consume a value of atomic base type (such as~\texttt{Int} or
        \texttt{Ptr}) exactly once, just evaluate it;
        % \item
        to consume a function value exactly once, apply it to one argument,
        and consume its result exactly once;
        % \item
        to consume a value of an algebraic datatype exactly once,
        pattern-match on it, and consume all its linear components exactly once.
        % For example, a linear pair (equivalent to $\tensor$) is consumed exactly
        % once if pattern-matched on \emph{and} both the first and second element are
        % consumed once.
%\end{itemize}
Thus, we can consume a linear resource by fully evaluating it -- and evaluation
happens in case expressions.

@@

In Core, cases are of the form $\ccase{e_s}{z~\{\ov{\rho_i \to e_i}\}}$,
where $e_s$ is the case \emph{scrutinee}, $z$ is the case \emph{binder}, and
$\ov{\rho_i \to e_i}$ are the case \emph{alternatives}, composed of a pattern
$\rho_i$ and of the corresponding expression $e_i$. 
%\begin{enumerate}
%\item
The case scrutinee is always evaluated to Weak Head Normal Form (WHNF) and
%
% \item
% evaluating an expression that is already in WHNF is a no-op, that is,
% no computation occurs; 
%
the case binder is an alias to the result of evaluating the scrutinee
to WHNF. Moreover, in Core, the alternative patterns are always exhaustive.
% i.e. there always exists
% a pattern that matches the WHNF of a value resulting from evaluating the
% scrutinee, where a pattern is either a wildcard that matches all expressions
% ($\_$), or a constructor and its linear and non-linear component binders
% ($K~\ov{x}\ov{y}$, with $\ov{x}$ as linearly-bound variables and $\ov{y}$ as
% unrestricted ones).

@@

We start with an
example of a program that constructs a tuple from linear resources then pattern
matches on it, then uses both linearly-bound variables from the tuple pattern
match. This is well-typed in Linear Haskell:
\begin{working}
\begin{lstlisting}
f6 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* c) -> c
f6 x y use = case (x,y) of z { (a,b) -> use a b }
\end{lstlisting}
\end{working}
What might be more surprising is that a similar program which discards the
pattern variables and instead uses the resources in the scrutinee is 
\emph{semantically} linear, despite not being accepted by Linear Haskell:
\begin{notyet}
\begin{lstlisting}
f7 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* c) -> c
f7 x y use = case (x,y) of z { (a,b) -> use x y }
\end{lstlisting}
\end{notyet}
We justify that \incode{f7} is linear by appealing to the fact that 
the tuple being scrutinized is already in WHNF and so evaluating it will
not consume either $x$ nor $y$. Even if the tuple was constructed with
two expressions using $x$ and $y$ respectively, no computation would happen
since we are not using $a$ or $b$ (thereby never forcing the arguments
of the tuple). However, if we did use $a$ in the case body, then $x$
would be unavailable (as opposed to \incode{y}) and \incode{f8} below must be rejected:
\begin{noway}
\begin{lstlisting}
f8 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* a *' $\lolli$ '* c) -> c
f8 x y use = case (x,y) of z { (a,b) -> use a x }
\end{lstlisting}
\end{noway}
This idea that $x$ and $a$ are mutually exclusive is the same behind let
bindings being mutually exclusive to the resources that define them.
By forcing the pattern variable we run the computations
defined in terms of the linear variables used for that constructor
argument. Otherwise, if we do not use the pattern variables, then we
do not run the computation and thus no resources are consumed.

A third (semantically well-typed, but rejected by Linear Haskell) option is to use the case binder $z$ instead of $a,b$ or $x,y$:
\begin{notyet}
\begin{lstlisting}
f9 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* c) -> c
f9 x y use = case (x,y) of z { (a,b) -> uncurry use z }
\end{lstlisting}
\end{notyet}
Again, $z$ is mutually exclusive with $(a,b)$ and with $(x,y)$, but at least one of
the three must occur to ensure the linear resources are consumed. In this
example, we can think that using $a$ entails using the resource $x$, $b$ the
resource $y$, and the case binder $z$ entails using both $a$ and $b$.

Dually, consider the scrutinee to be an expression that's not in WHNF, s.t.
evaluating it to WHNF will require performing computation and thus consume linear
resources (\incode{f10} below is well-typed in our system but rejected
by Linear Haskell):
\begin{notyet}
\begin{lstlisting}
f10 :: a *' $\lolli$ '* b *' $\lolli$ '* (a *' $\lolli$ '* b *' $\lolli$ '* (c,d)) -> (c,d)
f10 x y use = case use x y of z { (a,b) -> z }
\end{lstlisting}
\end{notyet}
Unlike when the scrutinee was in WHNF, we can no longer use $x,y$ in the case
alternatives, but we must still use either the case binder $z$ or the linear
pattern variables $a,b$. %  e.g. it would be quite disastrous if any of the following typechecked:
% \begin{noway}
% \begin{lstlisting}
% doubleFree :: Ptr *' $\lolli$ '* (Ptr *' $\lolli$ '* Result) -> Result
% doubleFree x free = case free x of z { Result v -> free x }
% \end{lstlisting}
% \end{noway}
% % Remover
% \begin{noway}
% \begin{lstlisting}
% leakPointer :: Ptr -> ()
% leakPointer x = case id x of z { _ -> () }
% \end{lstlisting}
% \end{noway}
%
%whereas this is fine:
%\begin{notyet}
%\begin{code}
%f16 :: Ptr -o (Ptr -o (Value,Int)) -> (Value,Int)
%f16 x free = case K (free x) of z { K y -> free x }
%\end{code}
%\end{notyet}
%
The result of evaluating the scrutinee must be consumed exactly once to guarantee
that the resources used in the scrutinee are fully consumed. %, or
                                %risk them being only ``almost''
                                %consumed.
For instance, take
\incode{use} in \incode{f10} to be the pairing function \incode{(,)}: it is not sufficient for \incode{use x y} to be
evaluated to WHNF to consume \incode{x} and \incode{y}. Otherwise, if all the resources were
considered to be fully consumed after the scrutinee were evaluated in a case
expression, we could simply ignore the pattern variables, potentially
discarding linear resources. In
short, if the scrutinee is not in WHNF we must either consume the case binder
or the linear components of the pattern.

On the other hand, if we match on a constructor without any linear components,
all linear resources used in the scrutinee must have been fully consumed in
that branch since the result of evaluating it can be consumed unrestrictedly
(by definition, a constructor application to unrestricted arguments is also
unrestricted).
%
Hence, in the branch for a constructor without linear fields, the case
binder can also be used unrestrictedly. For example, the program in
Figure~\ref{fig:nolin} is semantically linear.
%
\begin{figure}[h]
 \begin{minipage}{0.5\textwidth}
\begin{notyet}
\begin{lstlisting}
f11 :: () *' $\lolli$ '* ()
f11 x = case x of z { () -> z <> z }
\end{lstlisting}
\end{notyet}
\vspace{-0.5cm}
\caption{No Linear Fields\label{fig:nolin}}
\end{minipage}%
\begin{minipage}{0.5\textwidth}
\begin{limitation}
\begin{lstlisting}
f12 :: a *' $\lolli$ '* a
f12 x = case K1 x of z { K2 -> x; K1 a -> x }
\end{lstlisting}
\end{limitation}
\vspace{-0.5cm}
\caption{Absurd Branches\label{fig:absurd}}
\end{minipage}
\end{figure}
%
A second example of an unrestricted pattern, where \incode{K2} has no fields
and \incode{K1} has one linear field, is shown in Figure~\ref{fig:absurd}.
The use of \incode{x} in the \incode{K2} branch is valid despite \incode{x}
also occurring in the scrutinee because this branch is never taken. In fact,
any arbitrary resource could be used in absurd branches and be understood as
linear (despite not being seen as such by our system).

Many of these examples might seem unrealistic due to the pattern match on a
known constructor, but we note that in between the transformations programs
undergo in an optimising compiler, many such programs naturally occur (e.g.,~if
the definition of a function is inlined in the scrutinee).
%
% However, the scrutinee before evaluation likely isn't in WHNF, thus scrutinee
% resources cannot directly occur in the case alternatives, since they might be
% potentially consumed when evaluating the scrutinee, the program leading up to
% this one wouldn't have been linear (even semantically).

@@

As for \emph{default} case alternatives, also known as \emph{wildcards}
(written $\_$): matching a wildcard doesn't provide new linearity information,
unlike a pattern with linear or unrestricted components, however, the
scrutinee is still evaluated to WHNF.
%
% so linearity is seen as before but without
% fully consuming the scrutinee linear resources (in non-linear patterns) nor
% binding new linear resources (in linear patterns).
%
Briefly, when matching a wildcard, if the scrutinee is already in WHNF,
programs that either use the resources from the scrutinee directly, or the case
binder in that alternative, can be seen as linear. If the scrutinee is not in
WHNF, we \emph{must} use the case binder, as it's the only way to linearly
consume the result of evaluating the scrutinee.

Finally, we discuss the special case of scrutinizing a
single variable $x$, e.g.:
$
\lambda x.~\ccase{x}{\_ \to x}
$. 
It might seem as though this function is linear:
%begin{itemize}
% \item
if the argument instantiating $x$ is in WHNF, then scrutinizing it is a no-op, and returning $x$
just returns the resource intact;
% \item
if the argument instantiating $x$ is not in WHNF, then scrutinizing it evaluates it to WHNF, and returning
$x$ returns the result of evaluating $x$ that still had to be consumed exactly
once.
%\end{itemize}
However, in practice, this program's linearity can only be determined by
further analysis on the non-strict evaluation strategy. If linear
function applications are $\beta$-reduced \emph{call-by-name}
(often desirable, as linear functions are known to use their argument exactly
once)
% and thus call-by-name reduction will definitely not duplicate computation
while the example function is considered linear, then such an application might
duplicate linear resources during evaluation. For instance:
\[
\begin{array}{l}
(\lambda x.~\ccase{x}{\_ \to x})~(free~x)
\Longrightarrow_{CBN}
\ccase{free~x}{\_ \to free~x}
\end{array}
\]
On the other hand, if linear function applications are reduced
\emph{call-by-need}, the above example would instead be semantically linear
because the result of freeing $x$ would be bound to some $y$ first.
%
The system we present in the next section does accept this function
scrutinizing a variable as linear, as our operational semantics are
call-by-need.
This interaction between evaluation and linearity is considered specifically in the
reverse binder swap transformation discussed in
Section~\ref{sec:optimisations-preserve-types-meta}.\todo{Figure out precisely reverse binder swap proof and reevaluate}

@@

\section*{Acknowledgements}

We thank Arnaud Spiwack for insightful discussions and Well-Typed for funding
this work.\todo{Do these have to be anonymized?}
